
This appendix documents the programmatic endpoints, parameters, and query structures utilized to interface with external geospatial and economic APIs.

\section{B.1 Google Earth Engine REST API}
Earth Engine operations are authenticated using OAuth 2.0 with credentials routed through Google Cloud Projects. The pipeline initializes a connection using the base URL \texttt{https://earthengine.googleapis.com/v1alpha}. The initialization method \texttt{ee.Initialize(project='tesis-500804')} specifies the billing quota and project boundaries. When querying reference image collections, the pipeline uses the endpoint \texttt{/projects/{project}/imageCollections/{imageCollectionId}/images}. For optical data retrieval, the harmonized Sentinel-2 collection is queried and filtered geographically and temporally:
\texttt{}`python
ee.ImageCollection("COPERNICUS/S2\_SR\_HARMONIZED")
  .filterBounds(geometry)
  .filterDate("2024-01-01", "2024-12-31")
\texttt{}`

\section{B.2 Copernicus CDSE Process API}
The Copernicus Data Space Ecosystem (CDSE) requires generating a client token using keycloak OAuth endpoints, which is then passed as an Authorization header in subsequent Process API requests. The authentication endpoint is located at \texttt{https://identity.dataspace.copernicus.eu/auth/realms/CDSE/protocol/openid-connect/token}. The request payload is dispatched as a POST request:
\texttt{}`json
{
  "client\_id": "YOUR\_CLIENT\_ID",
  "client\_secret": "YOUR\_CLIENT\_SECRET",
  "grant\_type": "client\_credentials"
}
\texttt{}`
Once the bearer token is acquired, requests are directed to the Sentinel Hub Process API at \texttt{https://sh.dataspace.copernicus.eu/api/v1/process}. The request bands format specifies the required multi-spectral channels: \texttt{["B02", "B03", "B04", "B08", "B11", "QA60"]}.

\section{B.3 World Bank Open Data API}
Socioeconomic indicator datasets are downloaded using direct HTTP requests, querying the country three-letter codes:
\begin{itemize}
  \item \textbf{Query URL}: \texttt{http://api.worldbank.org/v2/country/{country}/indicator/{indicator}}
  \item \textbf{GDP Indicator Code}: \texttt{NY.GDP.MKTP.KD.ZG} (GDP growth, annual \%)
  \item \textbf{Default Output Format}: \texttt{json}

\end{itemize}
\section{B.4 NASA Earthdata Search \& Common Metadata Repository (CMR) API}
The pipeline queries NASA's Common Metadata Repository (CMR) API to identify and locate Suomi-NPP VIIRS Day/Night Band (DNB) monthly composites.
\begin{itemize}
  \item \textbf{Base Query URL}: \texttt{https://cmr.earthdata.nasa.gov/search/granules.json}
  \item \textbf{Parameters}: \texttt{short\_name=VNP46A1}, \texttt{temporal=2021-01-01T00:00:00Z,2026-05-31T23:59:59Z}, and bounding box filters.
  \item \textbf{Authentication}: Requires a Bearer token generated using NASA Earthdata login credentials.

\end{itemize}
\section{B.5 UN COMTRADE API}
The pipeline queries the UN COMTRADE database to fetch bilateral trade volumes for mineral commodity codes.
\begin{itemize}
  \item \textbf{Base Query URL}: \texttt{https://comtradeapi.un.org/data/v1/get/C/A/HS}
  \item \textbf{Commodity Codes}: \texttt{282520} (Lithium Oxide and Hydroxide), \texttt{7108} (Gold).
  \item \textbf{Authentication}: Requires a subscription key passed in the \texttt{Ocp-Apim-Subscription-Key} HTTP header.

\end{itemize}